---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Coffee Roasting Data Analysis

Roasts - One run

Timesteps - One step in time during a run

Events - Special steps in time during a run

Idea will be to have the following tables

-   `curves` (each row is a time step, each column is a input setting or output measurement, each value is the setting/measurement value at that step in time)

-   `events` (`event_type`, `roast_id`,

-   `parameters` (`roast_id`, `bean_name`, `batch_size`, `batch_yield`, `preheat_temp`

-   `metrics`

```{r}
library(tidyverse)
```

## Read in data

```{r}
paths <- list.files("data/my-roasts/")
D_roasts <- paths |>
  set_names(basename) |> 
  map(\(path) read_csv(str_c("data/my-roasts/",path),skip = 19)) |>
  list_rbind(names_to = "file") |>
  mutate(
    date = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 1),
    id = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 2),
    country = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 3),
    name = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 4),
    process = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 5),
    charge_weight = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 6),
    yield_weight = str_extract(file,"(\\d{4}-\\d{2}-\\d{2})_(\\d*)_(\\D*)_(\\D*)_(\\D*)_(\\d*)g_(\\d*)g.csv", group = 7),
  )
```

## Data Viz

```{r}
distinct(D_roasts, name, id)
```

```{r}
#Probe, IBTS, and ROR curves
D_roasts |> 
  filter(
    id == 1
  ) |> 
  ggplot() +
  geom_line(aes(x=Time, y = `Bean Probe Temp`)) +
  geom_line(aes(x=Time, y = `IBTS Temp`), color = "skyblue") +
  geom_line(aes(x=Time, y = `Bean Probe ROR`*10), color = "tan") +
  scale_y_continuous(limits = c(0,300)) +
  theme_minimal()
```

```{r}
#Bean probe tempreature curves
D_roasts |> 
  filter(
    id == 15 | id == 13 | id == 16
  ) |> 
  ggplot(aes(x=Time, y = `Bean Probe Temp`, color = id)) +
  geom_line(alpha=0.7) +
  scale_y_continuous(breaks = seq(0,220, by=10)) +
  theme_minimal()
  #labs(title = "El Salvador Atiquizaya Castillo")
```

```{r}
#IBTS tempreature curves + targets
D_roasts |> 
  # filter(
  #   id == 15 | id == 13 | id == 11
  # ) |> 
ggplot() +
  geom_rect(aes(xmin=hm("6:45"),xmax=hm("9:05"),ymin=198,ymax=200),
            fill="grey") +
  geom_rect(aes(xmin=hm("8:30"),xmax=hm("10:30"),ymin=205,ymax=209),
            fill="tan") +
  geom_hline(yintercept = 205, linetype="dotted", color="tan") +
  geom_rect(aes(xmin=hm("10:30"),xmax=hm("11:30"),ymin=209,ymax=212),
            fill="peru") +
  geom_hline(yintercept = 209, linetype="dotted", color="peru") +
  geom_rect(aes(xmin=hm("11:30"),xmax=hm("12:30"),ymin=212,ymax=218),
            fill="saddlebrown") +
  geom_hline(yintercept = 212, linetype="dotted", color="saddlebrown") +
  geom_line(aes(x=Time, y = `IBTS Temp`, color = id),alpha=0.7) +
  theme_minimal() +
  facet_wrap(~name)
  #labs(title = "El Salvador Atiquizaya Castillo - Infrared Temperature Sensor Curves")
```

```{r}
#Probe, IBTS, and ROR curves for a single roast id
D_roasts |> 
  filter(
    id == 15 | id == 13 | id == 16
  ) |> 
  ggplot() +
  geom_line(aes(x=Time, y = `Bean Probe Temp`)) +
  geom_line(aes(x=Time, y = `IBTS Temp`), color = "skyblue") +
  geom_line(aes(x=Time, y = `Bean Probe ROR`*10), color = "tan") +
  scale_y_continuous(limits = c(0,300)) +
  #labs(title = 'Roasts of El Salvador Atiquizaya Castillo')+
  ylab("")+
  theme_minimal() +
  facet_wrap(~id)
```

```{r}
D_roasts |> 
  ggplot() +
  geom_rect(aes(xmin=hm("6:45"),xmax=hm("9:05"),ymin=198,ymax=200),
            fill="grey") +
  geom_rect(aes(xmin=hm("8:30"),xmax=hm("10:30"),ymin=205,ymax=209),
            fill="tan") +
  geom_hline(yintercept = 205, linetype="dotted", color="tan") +
  geom_rect(aes(xmin=hm("10:30"),xmax=hm("11:30"),ymin=209,ymax=212),
            fill="peru") +
  geom_hline(yintercept = 209, linetype="dotted", color="peru") +
  geom_rect(aes(xmin=hm("11:30"),xmax=hm("12:30"),ymin=212,ymax=218),
            fill="saddlebrown") +
  geom_hline(yintercept = 212, linetype="dotted", color="saddlebrown") +
  geom_line(aes(x=Time, y = `Bean Probe Temp`)) +
  geom_line(aes(x=Time, y = `IBTS Temp`), color = "skyblue") +
  geom_line(aes(x=Time, y = `Bean Probe ROR`*10), color = "tan") +
  scale_y_continuous(limits = c(0,300)) +
  #labs(title = 'Roasts of Ethiopia Dry Process Korate')+
  theme_minimal() +
  facet_wrap(~id)
```

TODO - It would be nice to create a `bean` column and see how events change for different roasts

```{r}
data.frame(w = c(350, 500, 750, 1000),
           t = c(220, 250, 280, 310),
           l = c(200, 230, 275, 310),
           u = c(240, 275, 310, 310)) |> 
  ggplot() +
  geom_line(aes(x = w, y=t)) +
  geom_line(aes(x = w, y=l)) +
  geom_line(aes(x = w, y=u)) +
  geom_vline(xintercept=450, linetype="dashed")
```
