# Royal Crown Jewels

Aillio touts Roast World as a place to share your roasts. But in practice, this is pretty useless. Even if the person entered in all the relevant data for their roast (rare), how do I know their roast went well? How do I know that they're even a competent enough to judge "went well"? If it did go well, is that because

For a while I found it to be less of a gold mine than a junk heap. That is, until I stumbled upon Royal Coffee's Crown Jewels, roasting guides for the bullet for their top of the line coffees, by Evan Gilman.

```{r}
library(tidyverse)
```

```{r}
paths <- list.files("data/royal-crown-jewels/")

countries <- c("Kenya","Ethiopia","Uganda","Honduras","Panama","Sumatra","Colombia","Ecuador","Peru","Sulawesi","Rwanda","Brazil","Papua New Guinea","Burundi","Guatemala","Costa Rica","Bolivia","Mexico","Timor Leste","Tanzania","Bali")

D_jewels <- paths |>
  set_names(basename) |> 
  map(\(path) read_csv(str_c("data/royal-crown-jewels/",path),skip = 19)) |>
  list_rbind(names_to = "file") |> 
  mutate(
    date = str_extract(file,"^\\d{4}-\\d{2}-\\d{2}") |> ymd(),
    id = str_extract(file,"CJO?\\d{4}"),
    country = str_extract(file, paste(countries, collapse = "|"))
  ) |> 
  group_by(file) |> 
  mutate(
    preheat_temp = first(`IBTS Temp`),
    finish_temp = last(`IBTS Temp`),
    finish_time = last(Time)
  ) |> 
  ungroup()
```

```{r}
D_jewels |> filter(
  is.na(country)
)
```

```{r}
D_jewels |> filter(
  is.na(id)
)
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `IBTS Temp`, group = file, color=date)) +
  geom_line(alpha=0.3) +
  theme_minimal() +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Bean Probe Temp`, group = file, color=date)) +
  geom_line(alpha=0.1) +
  theme_minimal() +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Bean Probe ROR`, group = file, color=date)) +
  geom_line(alpha=0.1) +
  scale_y_continuous(limits = c(0,30)) +
  theme_minimal() +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Power Setting`, group=file)) +
  geom_area(alpha=0.1, position = "identity")+
  #scale_y_continuous(limits = c(0,30)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Fan Setting`, group=file)) +
  geom_area(alpha=0.1, position = "identity") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Drum Setting`, group=file)) +
  geom_area(alpha=0.1, position="identity")+
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~factor(year(date)))
```

TODO - End Temperature (y) vs End Time (x) faceted by year AND/OR colored by process method/origin/cultivar. Could also add in Yellowing Temperature and Yellowing Time as well as FC Temperature and FC Time to the same plot. All temperatures IBTS.

```{r}
D_jewels |> 
  ggplot(aes(x=date,y=preheat_temp)) +
  geom_point(size=0.7) +
  scale_y_continuous(breaks=c(200,210,220,230,240,250,260))+
  labs(title="Preheat temps over time")
```

```{r}
D_jewels |> 
  ggplot(aes(x=date,y=finish_time)) +
  geom_point(size=0.7) +
  labs(title="Finish times")
```

```{r}
D_jewels |> 
  ggplot(aes(x=date,y=finish_temp)) +
  geom_point(size=0.7) +
  labs(title="Finish temperatures")
```

```{r}
D_jewels |> 
  ggplot(aes(x=finish_time,y=finish_temp, group = file, color=date)) +
  geom_point(size=0.7) +
  geom_rug() +
  scale_y_continuous(limits = c(195,210)) +
  facet_wrap(~factor(year(date)))
```

TODO - Process Yellowing + First Crack Info from files. Show plots of times, temperatures of yellowing and first crack, as well as plots of times of percentages spent in each roast phase. (High value, can inform my roasting)

```{r}
D_jewels |> 
  ggplot(aes(x=finish_time, y=preheat_temp))+
  geom_point(size=0.7) +
  labs(title="Nice negative correlation between preheat temp and finish time")
```
