# Royal Crown Jewels

Aillio touts Roast World as a place to share your roasts. But in practice, this is pretty useless. Even if the person entered in all the relevant data for their roast (rare), how do I know their roast went well? How do I know that they're even a competent enough to judge "went well"? If it did go well, is that because

For a while I found it to be less of a gold mine than a junk heap. That is, until I stumbled upon Royal Coffee's Crown Jewels, roasting guides for the bullet for their top of the line coffees, by Evan Gilman.

```{r}
library(tidyverse)
library(GGally)
library(umap)
library(ggiraph)
library(patchwork)
library(stringi)
```

## 1. Wrangling

```{r}
#Get roast curves

paths <- list.files("data/royal-crown-jewels/")

D_curves <- paths |> 
  set_names(basename) |> 
  map(\(path) read_csv(str_c("data/royal-crown-jewels/",path),
                       skip = 19,
                       col_types = cols(Time = col_character())
                       )
      ) |>
  list_rbind(names_to = "file") |>
  janitor::clean_names() |>
  mutate(
    time = ms(time),
    date = str_extract(file,"^\\d{4}-\\d{2}-\\d{2}") |> ymd(),
    year = year(date),
    name = str_extract(file,"^\\d{4}-\\d{2}-\\d{2}_(.+).csv$", group = 1) |> 
      enc2utf8() |> 
      stri_trans_nfc() |> 
      str_trim(),
    id = str_extract(file,"CJO?\\d{4}")
  )
```

```{r}
#Get Yellow + First Crack time markers from CSVs

paths <- list.files("data/royal-crown-jewels/")

D_roasts <- paths |> 
  set_names(basename) |> 
  map(\(path) read_csv(str_c("data/royal-crown-jewels/",path),
                       n_max=19,
                       col_types = cols(c3 = col_character())
                       )
      ) |>
  list_rbind(names_to = "file") |>
  group_by(file) |> 
  mutate(
    time_yellow = nth(c3, 16) |> ms(),
    time_fc = nth(c3, 9) |> ms()
  ) |> 
  ungroup() |> 
  distinct(file, time_yellow, time_fc) 

#Get event data from roast curves

D_roasts <- D_curves |> 
  right_join(D_roasts,by = join_by(file)) |>
  group_by(file) |> 
  mutate(
    temp_ibts_start = first(ibts_temp),
    temp_ibts_yellow = ibts_temp[time == time_yellow],
    temp_ibts_fc = ibts_temp[time == time_fc],
    temp_ibts_end = last(ibts_temp),
    temp_bean_probe_start = first(bean_probe_temp),
    temp_bean_probe_turning = min(bean_probe_temp),
    temp_bean_probe_yellow = bean_probe_temp[time == time_yellow],
    temp_bean_probe_fc = bean_probe_temp[time == time_fc],
    temp_bean_probe_end = last(bean_probe_temp),
    time_turning = time[bean_probe_temp == temp_bean_probe_turning][1],
    ror_bean_probe_max = max(bean_probe_ror),
    time_max_ror = time[bean_probe_ror == ror_bean_probe_max][1],
    duration_drying = time_yellow,
    duration_maillard = time_fc - time_yellow,
    duration_development = last(time) - time_fc,
    duration_total = last(time),
    pct_drying = duration_drying / duration_total,
    pct_maillard = duration_maillard /duration_total,
    pct_development = duration_development / duration_total
  ) |> 
  ungroup() |> 
  select(-c(timeline:fan_setting)) |> 
  distinct()


#Get origin from file

countries <- c("Kenya","Ethiopia","Uganda","Honduras","Panama","Sumatra","Colombia","Ecuador","Peru","Sulawesi","Rwanda","Brazil","Papua New Guinea","Burundi","Guatemala","Costa Rica","Bolivia","Mexico","Timor Leste","Tanzania","Bali")


D_roasts <- D_roasts |> 
  mutate(
    country = str_extract(file, paste(countries, collapse = "|"))
  )
```

```{r}
#Get data on physical bean characteristics
D_bean_char <- read_csv('data/crown_jewels_urls_loss.csv') |> 
  mutate(
    name = name |> enc2utf8() |> stri_trans_nfc()
  ) |> 
  select(-c(id,bean,url,pdf))

# anti.join <- D_bean_char |> 
#   anti_join(D_roasts, by = join_by(name_from_csv == name_from_file))

D_roasts <- D_roasts |>
  right_join(D_bean_char, by = join_by(name))
```

⚠️ TODO - Learn more about processing. Seems like a leaky abstraction.

```{r}
# mutate(
#   process_simplified = case_when(
#     str_detect(file,
#                regex("Decaf", ignore_case = TRUE)) ~ "decaf",
#     str_detect(file,
#                regex("Anaerobic", ignore_case = TRUE)) ~ "anaerobic",
#     str_detect(file,
#                regex("Honey", ignore_case = TRUE)) ~ "honey",
#     str_detect(file,
#                regex("Hull", ignore_case = TRUE)) ~ "wet hulled",
#     str_detect(file,
#                regex("Monsoon", ignore_case = TRUE)) ~ "monsoon",
#     str_detect(file,
#                regex("Wet|Washed", ignore_case = TRUE)) ~ "washed",
#     str_detect(file,
#                regex("Dry|Natural", ignore_case = TRUE)) ~ "dry",
#     TRUE ~ NA_character_
#   )
# ) |> 
```

## 2. Summary Stats

Table stakes: weight loss (12.5% ± 1%) , drying ratio (42% ± 3%), maillard ratio (40% ± 5%), development ratio (17% ± 4%) , end temperatures.

```{r}
D_roasts |> 
  filter(year >= 2023) |> 
  summarise(
    mean_dry = mean(pct_drying),
    sd_dry = sd(pct_drying),
    mean_mai = mean(pct_maillard),
    sd_mai = sd(pct_maillard),
    mean_dev = mean(pct_development),
    sd_dev = sd(pct_development),
  )
```

```{r}
D_roasts |> count(country)
```

```{r}
D_roasts |>
  summarise(
    mean_p_free = mean(density_free_settled, na.rm=TRUE),
    sd_p_free = sd(density_free_settled, na.rm=TRUE),
    mean_p_sinar = mean(density_sinar, na.rm=TRUE),
    sd_p_sinar = sd(density_sinar, na.rm=TRUE),
    mean_pct_moisture = mean(moisture_pct, na.rm=TRUE),
    sd_pct_moisture = sd(moisture_pct, na.rm=TRUE),
    mean_water_activity = mean(water_activity, na.rm=TRUE),
    sd_water_activity = sd(water_activity, na.rm=TRUE)
  )
```

## 3. EDA

### a. Sensor Curves by Year

```{r}
D_curves |> 
  ggplot(aes(x=time, y = ibts_temp, group = file)) +
  # geom_rect(aes(xmin=ms("6:45"),xmax=ms("9:05"),ymin=198,ymax=200),
  #           fill="grey") +
  # geom_rect(aes(xmin=ms("8:30"),xmax=ms("10:30"),ymin=205,ymax=209),
  #           fill="tan") +
  # geom_hline(yintercept = 205, linetype="dotted", color="tan") +
  # geom_rect(aes(xmin=ms("10:30"),xmax=ms("11:30"),ymin=209,ymax=212),
  #           fill="peru") +
  # geom_hline(yintercept = 209, linetype="dotted", color="peru") +
  # geom_rect(aes(xmin=ms("11:30"),xmax=ms("12:30"),ymin=212,ymax=218),
  #           fill="saddlebrown") +
  # geom_hline(yintercept = 212, linetype="dotted", color="saddlebrown") +
  geom_line(aes(color=date),alpha=0.3) +
  theme_minimal() +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  facet_wrap(~year) +
  labs(title = "IBTS Curves by Year")
```

```{r}
D_curves |> 
  ggplot(aes(x=time, y = bean_probe_temp, group = file, color=date)) +
  geom_line(alpha=0.1) +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  theme_minimal() +
  facet_wrap(~year) +
  labs(title = "BT Curves by Year")
```

```{r}
D_curves |> 
  ggplot(aes(x=time, y = bean_probe_ror, group = file, color=date)) +
  geom_line(alpha=0.1) +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  scale_y_continuous(limits = c(0,30)) +
  theme_minimal() +
  facet_wrap(~year) +
  labs(title = "BT ROR Curves by Year")
```

### b. Profile Curves by Year

⚠️TODO - Normalize these by year.

```{r}
D_curves |> 
  group_by(year, time, power_setting) |> 
  summarise(n = n_distinct(file), .groups = 'drop') |> 
  ungroup() |>
  ggplot(aes(x=time, y = power_setting, alpha = n/143)) +
  geom_tile(fill="red") +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~year) +
  labs(title = "Power Setting Profiles by Year")
```

```{r}
D_curves |> 
  group_by(year, time, fan_setting) |> 
  summarise(n = n_distinct(file), .groups = 'drop') |> 
  ungroup() |>
  ggplot(aes(x=time, y = fan_setting, alpha = n/143)) +
  geom_tile(fill="red") +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~year) +
  labs(title = "Fan Setting Profiles by Year")
```

```{r}
D_curves |> 
  group_by(year, time, drum_setting) |> 
  summarise(n = n_distinct(file), .groups = 'drop') |> 
  ungroup() |>
  ggplot(aes(x=time, y = drum_setting, alpha = n/143)) +
  geom_tile(fill="red") +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~year) +
  labs(title = "Drum Setting Profiles by Year")
```

```{r}
power.profiles <- D_curves |> 
  filter(year >= 2023) |> 
  pivot_longer(
    cols = c(power_setting,drum_setting,fan_setting),
    names_to = "control",
    values_to = "setting",
    names_pattern = "(.+)_setting$"
  ) |>
  group_by(control, time, setting) |> 
  summarise(n = n_distinct(file), .groups = "drop") |>
  ungroup()

ggplot() +
  geom_tile(data=power.profiles, aes(x = time, y = setting, fill = n)) +
  geom_histogram(data=D_roasts |> filter(year(date) >= 2023), 
             aes(x=time_yellow,
                 y = after_stat(count / 8)),
             alpha=0.3,
             bins = 30,
             fill="gold") +
  geom_histogram(data=D_roasts |> filter(year(date) >= 2023), 
             aes(x=time_fc,
                 y = after_stat(count / 8)),
             alpha=0.3,
             bins = 30) +
  geom_histogram(data=D_roasts |> filter(year(date) >= 2023), 
             aes(x=duration_total,
                 y = after_stat(count / 8)),
             alpha=0.3,
             bins = 30,
             fill="skyblue") +
  facet_wrap(~control, ncol=1) + 
  scale_x_time(limits = c(ms("0:00"),ms("13:00")),
               breaks = c(ms("0:00"),ms("1:00"),ms("2:00"),ms("3:00"),ms("4:00"),ms("5:00"),ms("6:00"),ms("7:00"),ms("8:00"),ms("9:00"),ms("10:00"),ms("11:00"),ms("12:00"),ms("13:00"))) +
  scale_y_continuous(breaks=0:9) +
  scale_fill_continuous(palette = c("white", "#FEE0D2", "#FC9272", "#DE2D26"))+
  theme_minimal() +
  labs(title="Settings Heatmaps for 2023 onward",
       subtitle = "Histograms show phase change time distributions")
```

⚠️IDEA - model roast phase (\< turning point, \< peak ROR, \< Yellowing, \< FC, \< End) as an ordinal variable (1, 2, 3, 4, 5)

### c. Event Temperatures by Year

```{r}
D_roasts |> 
  filter(year >= 2023) |> 
  summarise(
    mean_ibts_preheat_temp = mean(temp_ibts_start),
    sd_ibts_preheat_temp = sd(temp_ibts_start),
    mean_probe_preheat_temp = mean(temp_bean_probe_start),
    sd_probe_preheat_temp = sd(temp_bean_probe_start),
  )
```

This is another great presentation graphic. Things that change versus things that stay the same.

```{r}
D_roasts |> 
  ggplot(aes(x=date)) +
  geom_smooth(aes(y = temp_ibts_yellow), color = "gold") +
  geom_smooth(aes(y = temp_ibts_fc), color = "peru") +
  geom_smooth(aes(y = temp_ibts_end), color="#008080") +
  geom_smooth(aes(y = temp_ibts_start), color="black") +
  geom_smooth(aes(y = temp_bean_probe_start), color="purple") +
  geom_point(aes(y = temp_ibts_yellow), color = "gold", alpha=0.7) +
  geom_point(aes(y = temp_ibts_fc), color = "peru", alpha=0.7) +
  geom_point(aes(y = temp_ibts_end), color = "#008080",alpha=0.7) +
  geom_point(aes(y = temp_ibts_start), color = "black",alpha=0.7) +
  geom_point(aes(y = temp_bean_probe_start), color="purple", alpha=0.7) +
  scale_y_continuous(limits = c(140,260), n.breaks = 10) +
  theme_minimal()+
  ylab("Temperature (ºC)") +
  labs(title = "Event Temperatures by Year",
       subtitle = "Preheat, Yellowing, First Crack, and End Temperatures")
```

### d. Phase Duration by Year

These are really cool. They show what remains constant (share of time spent in roast phases) versus what changes (total roast time) as an experienced roaster gets acquainted with the Bullet.

Seems like phase shares are targets? **Would be curious to know if there's any structure to the residuals or if these targets are universal across greens?**

```{r}
D_roasts |> 
  pivot_longer(
    cols = c(pct_drying, pct_maillard, pct_development),
    names_to = "phase",
    values_to = "pct"
  ) |> 
  mutate(
    phase = factor(phase, levels = c("pct_development", "pct_maillard", "pct_drying"))
  ) |> 
  ggplot(aes(x=file, y = pct, fill = phase)) +
  geom_col() +
  scale_fill_manual(values = c("firebrick","gold","forestgreen")) +
  scale_x_discrete(labels = NULL) +
  labs(title="Phase Share by Roast Number")
```

```{r}
D_roasts |> 
  pivot_longer(
    cols = c(duration_drying, duration_maillard, duration_development),
    names_to = "phase",
    values_to = "duration"
  ) |> 
  mutate(
    phase = factor(phase, levels = c("duration_development", "duration_maillard", "duration_drying"))
  ) |> 
  ggplot(aes(x=file, y = duration, fill = phase)) +
  geom_col() +
  scale_fill_manual(values = c("firebrick","gold","forestgreen")) +
  scale_x_discrete(labels = NULL) +
  scale_y_time() +
  labs(title="Phase Duration by Roast Number")
```

### e. Univariate and Bivariate Distributions of Scalar Parameters

```{r}
D_roasts |> 
  pivot_longer(
    cols = c(pct_drying, pct_maillard, pct_development),
    names_to = "phase"
  ) |> 
  ggplot(aes(x=value)) +
  scale_x_continuous(limits = c(0,0.55),
                     n.breaks=10) +
  geom_histogram(bins=18) +
  facet_wrap(~phase)
```

```{r}
D_roasts |> 
  mutate(
    across(where(is.period), time_length)
  ) |> 
  summarise(
    mean_duration_dry = mean(duration_drying, na.rm=TRUE),
    sd_duration_dry = sd(duration_drying, na.rm=TRUE),
    mean_duration_mai = mean(duration_maillard, na.rm=TRUE),
    sd_duraiton_mai = sd(duration_maillard, na.rm=TRUE),
    mean_duration_dev = mean(duration_development, na.rm=TRUE),
    sd_duration_dev = sd(duration_development, na.rm=TRUE)
  )
```

```{r}
D_roasts |> 
  pivot_longer(
    cols = c(duration_drying, duration_maillard, duration_development),
    names_to = "phase"
  ) |> 
  ggplot(aes(x=value)) +
  scale_x_time() +
  geom_histogram(bins=20) +
  facet_wrap(~phase)
```

```{r}
#Pairwise Distributions + Correlations
p_to_stars <- function(p) {
  if (is.na(p)) ""
  else if (p < 0.001) "***"
  else if (p < 0.01)  "**"
  else if (p < 0.05)  "*"
  else ""
}

panel_cor_bg <- function(data, mapping, ...) {
  x <- eval_data_col(data, mapping$x)
  y <- eval_data_col(data, mapping$y)
  
  ct <- suppressWarnings(cor.test(x, y, use = "complete.obs"))
  r  <- unname(ct$estimate)
  p  <- ct$p.value
  stars <- p_to_stars(p)

  ggplot(data.frame(r), aes(x = 0, y = 0, fill = r)) +
    geom_tile(width = 1, height = 1) +
    scale_fill_gradient2(
      low = "#b2182b",
      mid = "white",
      high = "#4d9221",
      midpoint = 0,
      limits = c(-1, 1)
    ) +
    geom_text(aes(label = sprintf("%.2f%s", r, stars)), size = 4) +
    theme_void() +
    theme(legend.position = "none")
}

quiet <- function(fun) {
  function(...) suppressWarnings(fun(...))
}
```

```{r}
D_roasts |> 
  select(-c(file,date,name,id,country)) |> 
  mutate(
    across(where(is.period), time_length)
  ) |> 
  drop_na() |> 
  ggpairs(
    lower = list(
      continuous = quiet(wrap("points", size = 0.7, alpha = 0.3))
    ),
    upper = list(
      continuous = quiet(panel_cor_bg)
    ),
    diag = list(
      continuous = quiet(wrap("densityDiag"))
    ),
    progress = FALSE
  ) 
```

### f. UMAP

```{r}
roasts.umap <- D_roasts |> 
  select(-c(file,date,name,id,country)) |> 
  mutate(
    across(where(is.period), time_length)
  ) |>  
  drop_na() |> 
  umap()
```

```{r}
plot(roasts.umap$layout)
```

```{r}
D_roasts |> 
  select(-c(file,date,name,id,country)) |> 
  mutate(
    across(where(is.period), time_length)
  ) |>  
  drop_na() |> 
  bind_cols(roasts.umap$layout) |> 
  rename(
    umap_x = `...31`,
    umap_y = `...32`
  ) |> 
  ggplot(aes(x=umap_x, y=umap_y, color=time_length(duration_total))) +
  geom_point(alpha = 0.7)
```

```{r}
p.umap <- D_roasts |> 
  select(-c(date,name,id,country)) |> 
  mutate(
    across(where(is.period), time_length)
  ) |>  
  drop_na() |> 
  bind_cols(roasts.umap$layout) |> 
  rename(
    umap_x = `...32`,
    umap_y = `...33`
  ) |> 
  ggplot(aes(x=umap_x, y=umap_y, color=time_length(duration_total))) +
  geom_point_interactive(alpha=0.7, aes(data_id = file))

p.curves <- D_curves |> 
  ggplot(aes(x=time, y = bean_probe_ror, group = file))+
  geom_line_interactive(alpha=0.1, aes(data_id = file)) +
  scale_x_time(limits = c(ms("0:00"),ms("15:00"))) +
  scale_y_continuous(limits = c(0,30)) +
  theme_minimal()

girafe(ggobj = p.umap + p.curves, width_svg = 10, height_svg = 5) %>%
  girafe_options(opts_hover(css = "stroke:red; stroke-width:2; stroke-opacity:1;"))
```

```{r}
D_roasts |> 
  select(-c(file,date,name,id,country)) |> 
  mutate(
    across(where(is.period), time_length)
  ) |>  
  drop_na() |> 
  bind_cols(roasts.umap$layout) |> 
  rename(
    umap_x = `...31`,
    umap_y = `...32`
  ) |>
  ggpairs(
    lower = list(
      continuous = wrap("points", size = 0.7, alpha = 0.3)
    ),
    upper = list(
      continuous = panel_cor_bg
    ),
    progress = FALSE
  ) 
```

### g. Moisture Loss

```{r}
D_roasts |> 
  ggplot(aes(x = loss_pct)) +
  geom_histogram(bins = 11)
```

```{r}
D_roasts |> 
  summarise(
    loss.mean = mean(loss_pct),
    loss.sd = sd(loss_pct)
  )
```

```{r}
D_roasts |> 
  summarise(
    excess.loss.mean = mean(loss_pct - moisture_pct, na.rm = TRUE),
    excess.loss.sd = sd(loss_pct - moisture_pct, na.rm = TRUE)
  )
```

```{r}
D_roasts |> 
  summarise(
    moisture.mean = mean(moisture_pct, na.rm = TRUE),
    moisture.sd = sd(moisture_pct, na.rm = TRUE)
  )
```

```{r}
D_roasts |> 
  ggplot(aes(x=moisture_pct, y=loss_pct)) +
  geom_abline(slope = 1, intercept = 0, alpha = 0.3) +
  geom_abline(slope = 1, intercept = 1, alpha = 0.3) +
  geom_abline(slope = 1, intercept = 2, alpha = 0.3) +
  geom_abline(slope = 1, intercept = 3, alpha = 0.3) +
  geom_abline(slope = 1, intercept = 4, alpha = 0.3) +
  geom_vline(xintercept = 10.64) +
  geom_hline(yintercept = 12.53) +
  geom_point(alpha=0.7) +
  geom_rug(alpha = 0.3)+
  scale_x_continuous(limits = c(8,16)) +
  scale_y_continuous(limits = c(8,16)) 
```

```{r}
D_roasts |> 
  ggplot(aes(x=loss_pct-moisture_pct)) +
  geom_histogram(bins=10)
```

TODO - t tests + distributions for all parameters against process method and origin

TODO - Bivariate scatters + correlation coeiffs for roast prameters against density, moisture, and water activity

TODO - Read roast logs and come up with testable hypotheses for understanding control variables (preheat, power, fan, drum)
