# Royal Crown Jewels

Aillio touts Roast World as a place to share your roasts. But in practice, this is pretty useless. Even if the person entered in all the relevant data for their roast (rare), how do I know their roast went well? How do I know that they're even a competent enough to judge "went well"? If it did go well, is that because

For a while I found it to be less of a gold mine than a junk heap. That is, until I stumbled upon Royal Coffee's Crown Jewels, roasting guides for the bullet for their top of the line coffees, by Evan Gilman.

```{r}
library(tidyverse)
```

```{r}
paths <- list.files("data/royal-crown-jewels/")

countries <- c("Kenya","Ethiopia","Uganda","Honduras","Panama","Sumatra","Colombia","Ecuador","Peru","Sulawesi","Rwanda","Brazil","Papua New Guinea","Burundi","Guatemala","Costa Rica","Bolivia","Mexico","Timor Leste","Tanzania","Bali")

D_jewels <- paths |>
  set_names(basename) |> 
  map(\(path) read_csv(str_c("data/royal-crown-jewels/",path),
                       skip = 19,
                       show_col_types = FALSE)) |>
  list_rbind(names_to = "file") |> 
  mutate(
    date = str_extract(file,"^\\d{4}-\\d{2}-\\d{2}") |> ymd(),
    id = str_extract(file,"CJO?\\d{4}"),
    country = str_extract(file, paste(countries, collapse = "|"))
  ) |> 
  group_by(file) |> 
  mutate(
    preheat_temp = first(`IBTS Temp`),
    finish_temp = last(`IBTS Temp`),
    finish_time = last(Time)
  ) |> 
  ungroup()
```

```{r}
D_events <- paths |> 
  set_names(basename) |> 
  map(\(path) read_csv(str_c("data/royal-crown-jewels/",path),
                       n_max=19,
                       show_col_types = FALSE)) |>
  list_rbind(names_to = "file") |>
  group_by(file) |> 
  mutate(
    yellow_time = nth(c3, 16) |> parse_time(),
    fc_time = nth(c3, 9) |> parse_time()
  ) |> 
  ungroup() |> 
  distinct(file, yellow_time, fc_time)
```

```{r}
D_jewels <- D_jewels |> 
  right_join(D_events,by = join_by(file)) |> 
  group_by(file) |> 
  mutate(
    fc_temp = `IBTS Temp`[Time == fc_time],
    yellow_temp = `IBTS Temp`[Time == yellow_time]
  ) |>
  ungroup()
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `IBTS Temp`, group = file, color=date)) +
  geom_line(alpha=0.3) +
  theme_minimal() +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Bean Probe Temp`, group = file, color=date)) +
  geom_line(alpha=0.1) +
  theme_minimal() +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Bean Probe ROR`, group = file, color=date)) +
  geom_line(alpha=0.1) +
  scale_y_continuous(limits = c(0,30)) +
  theme_minimal() +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Power Setting`, group=file)) +
  geom_area(alpha=0.1, position = "identity")+
  #scale_y_continuous(limits = c(0,30)) +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Fan Setting`, group=file)) +
  geom_area(alpha=0.1, position = "identity") +
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  ggplot(aes(x=Time, y = `Drum Setting`, group=file)) +
  geom_area(alpha=0.1, position="identity")+
  theme_minimal() +
  scale_y_continuous(limits = c(0,9), breaks = c(0,1,2,3,4,5,6,7,8,9)) +
  facet_wrap(~factor(year(date)))
```

```{r}
D_jewels |> 
  distinct(file, date, preheat_temp) |> 
  ggplot(aes(x=date,y=preheat_temp)) +
  geom_point(size=1, alpha=0.6) +
  scale_y_continuous(breaks=c(200,210,220,230,240,250,260))+
  labs(title="Charge Temperature v.s. Roast Date")
```

```{r}
D_jewels |>
  distinct(file, date, finish_time) |> 
  ggplot(aes(x=date,y=finish_time)) +
  geom_point(size=1, alpha=0.6) +
  labs(title="Drop Time v.s. Roast Date")
```

```{r}
D_jewels |>
  distinct(file, date, finish_temp) |>
  ggplot(aes(x=date,y=finish_temp)) +
  geom_point(size=1, alpha=0.6) +
  scale_y_continuous(limits=c(195,210))+
  labs(title="Drop temperature v.s. Roast Date")
```

```{r}
D_jewels |> 
  distinct(file, date, fc_temp)|>
  ggplot(aes(x=date,y=fc_temp)) +
  geom_point(size=1, alpha=0.6) +
  #scale_y_continuous(limits=c(195,210))+
  labs(title="First Crack Temperature v.s. Roast Date")
```

```{r}
D_jewels |> 
  distinct(file, date, fc_time)|>
  ggplot(aes(x=date,y=fc_time)) +
  geom_point(size=1, alpha=0.6) +
  #scale_y_continuous(limits=c(195,210))+
  labs(title="First Crack Time v.s. Roast Date")
```

```{r}
D_jewels |> 
  distinct(file, date, yellow_temp)|>
  ggplot(aes(x=date,y=yellow_temp)) +
  geom_point(size=1, alpha=0.6) +
  #scale_y_continuous(limits=c(195,210))+
  labs(title="Yellowing Temperature v.s. Roast Date")
```

```{r}
D_jewels |> 
  distinct(file, date, yellow_time)|>
  ggplot(aes(x=date,y=yellow_time)) +
  geom_point(size=1, alpha=0.6) +
  #scale_y_continuous(limits=c(195,210))+
  labs(title="Yellowing Time v.s. Roast Date")
```

```{r}
D_jewels |>
  distinct(file, date, finish_temp, finish_time) |>
  ggplot(aes(x=finish_time,y=finish_temp, group = file, color=date)) +
  geom_point(size=0.7) +
  geom_rug() +
  scale_y_continuous(limits = c(195,210)) +
  facet_wrap(~factor(year(date))) +
  labs(title = "Drop Time and Temperature")
```

```{r}
D_jewels |>
  distinct(file, date, fc_temp, fc_time) |>
  ggplot(aes(x=fc_time,y=fc_temp, group = file, color=date)) +
  geom_point(size=0.7) +
  geom_rug() +
  #scale_y_continuous(limits = c(195,210)) +
  facet_wrap(~factor(year(date))) +
  labs(title = "First Crack Time and Temperature")
```

```{r}
D_jewels |>
  distinct(file, date, yellow_temp, yellow_time) |>
  ggplot(aes(x=yellow_time,y=yellow_temp, group = file, color=date)) +
  geom_point(size=0.7) +
  geom_rug() +
  #scale_y_continuous(limits = c(195,210)) +
  facet_wrap(~factor(year(date))) +
  labs(title = "Yellowing Time and Temperature")
```

TODO - Plots of times of percentages spent in each roast phase. (High value, can inform my roasting)

```{r}
D_jewels |>
  distinct(file, preheat_temp, fc_temp) |>
  ggplot(aes(x=fc_temp, y=preheat_temp))+
  geom_point(size=0.7) +
  geom_smooth(method="lm")+
  labs(title="Nice negative correlation between preheat temp and finish time")
```
